### 2018-08-02 doc/2019-01-poster/poster.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Jacobs Landscape Poster
% LaTeX Template
% Version 1.1 (14/06/14)
%
% Created by:
% Computational Physics and Biophysics Group, Jacobs University
% https://teamwork.jacobs-university.de:8443/confluence/display/CoPandBiG/LaTeX+Poster
% 
% Further modified by:
% Nathaniel Johnston (nathaniel@njohnston.ca)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[final]{beamer}

\usepackage[scale=0.9]{beamerposter} % Use the beamerposter package for laying out the poster
\usetheme{confposter} % Use the confposter theme supplied with this template

\setbeamercolor{block title}{fg=dblue!80,bg=white} % Colors of the block titles
\setbeamercolor{block body}{fg=black,bg=white} % Colors of the body of blocks
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks
\setbeamertemplate{caption}[numbered]

% Many more colors are available for use in beamerthemeconfposter.sty

%-----------------------------------------------------------
% Define the column widths and overall poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% In this template, the separation width chosen is 0.024 of the paper width and a 4-column layout
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. (1-(4+1)*0.024)/4 = 0.22
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. 
% (1-(3+1)*0.025)/3 = 0.3
% Set twocolwid to be (2*onecolwid)+sepwid = 0.464
% Set threecolwid to be (3*onecolwid)+2*sepwid = 0.708
\newcommand{\Cyclus}{\textsc{Cyclus}\xspace}%

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{33.1in} % A0 width: 46.8in
\setlength{\paperheight}{46.8in} % A0 height: 33.1in
\setlength{\textwidth}{33.1in} % A0 width: 46.8in
\setlength{\textheight}{46.8in} % A0 height: 33.1in
\setlength{\sepwid}{0.025\paperwidth} % Separation width (white space) between columns
\setlength{\onecolwid}{0.3\paperwidth} % Width of one column
\setlength{\twocolwid}{0.625\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.95\paperwidth} % Width of three columns
\setlength{\topmargin}{-1in} % Reduce the top margin size
%-----------------------------------------------------------

\usepackage{graphicx}  % Required for including images
\usepackage{hyperref}
\usepackage{cite}

\usepackage{tabularx}
\newcolumntype{b}{X}
\newcolumntype{s}{>{\hsize=.5\hsize}X}
\newcolumntype{m}{>{\hsize=.75\hsize}X}
\newcolumntype{z}{>{\hsize=.65\hsize}X}

\usepackage{booktabs} % Top and bottom rules for tables
\usepackage{xspace}

\usepackage{tikz}
\usepackage{chronology}
\usetikzlibrary{arrows.meta}

\usetikzlibrary{positioning, arrows, decorations, shapes, calc }
% Define block styles
\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]

\tikzstyle{const} = [rectangle, draw, text centered, fill=orange!20]
\tikzstyle{data} = [rectangle, draw, text centered, fill=green!20]

\tikzstyle{block} = [rectangle, draw, text centered, fill=blue!20]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=6em,
minimum height=2em]

\usetikzlibrary{shapes.multipart}
\usetikzlibrary{positioning}

%-----------------------------------------------------------
% KNOWN ISSUE IN TIKZ MATH: tiny big symbols, like sum.
%-----------------------------------------------------------

% declare `cmex` to be arbitrary scalable
\DeclareFontShape{OMX}{cmex}{m}{n}{
  <-7.5> cmex7
  <7.5-8.5> cmex8
  <8.5-9.5> cmex9
  <9.5-> cmex10
}{}

\SetSymbolFont{largesymbols}{normal}{OMX}{cmex}{m}{n}
\SetSymbolFont{largesymbols}{bold}  {OMX}{cmex}{m}{n}

\usepackage{amssymb}
\usepackage{pifont}
\usepackage{xcolor}
\newcommand{\greencheck}{{\color{green}\checkmark}}
\newcommand{\xmark}{{\color{red}\ding{55}}}

\setbeamertemplate{bibliography item}[text]

\makeatletter
\def\beamer@andinst{\\[0em]}
\makeatother

%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------

\title{Dynamic Transition Analysis with TIMES}

\author{
       % \textbf{ENERGY ANALYSIS DIVISION}\\
       % \textbf{I\textsuperscript{2}CNER Initiative on Challenges in Energy Assessment and Energy Transitions}\\
        Anshuman Chaube\inst{1}, Andrew Chapman\inst{2}, James Stubbins\inst{1}, \textbf{Kathryn Huff}\inst{1}}
\institute[shortinst]{ \inst{1} University of Illinois at Urbana-Champaign, Department of Nuclear, Plasma, and Radiological Engineering, Urbana, USA \and \inst{2} International Institute for Carbon Neutral Energy Research (I$^2$CNER), Kyushu University, Fukuoka, Japan}
%----------------------------------------------------------------------------------------

\begin{document}

\addtobeamertemplate{block end}{}{\vspace*{2ex}} % White space under blocks
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}} % White space under highlighted (alert) blocks

\setlength{\belowcaptionskip}{2ex} % White space under figures
\setlength\belowdisplayshortskip{2ex} % White space under equations

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

\begin{columns}[t,totalwidth=\threecolwid] % The whole poster consists of three major columns, the second of which is split into two columns twice - the [t] option aligns each column's content to the top

%=======================================================
% FIRST COLUMN BEGINS
%=========================================================

\begin{column}{\twocolwid} % The first two columns, two cols wide
\begin{columns}[t,totalwidth=\twocolwid] % Split up the first column into two, 
\begin{column}{\onecolwid} % The first column, one col wide

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

\begin{alertblock}{Objectives}

\begin{description}
{\large 
        \item[\textbf{Division:}] Energy Analysis Division
        \item[\textbf{Project:}] I\textsuperscript{2}CNER Initiative on Challenges in Energy Assessment and Energy Transitions
        \item[\textbf{Objective:}] Evaluate potential impact of novel energy technologies within Japan's energy system.
\item[\textbf{Milestones:}] ~\\}
        \begin{itemize}
        \item Minimize carbon emissions within realistic constraints.
        \item Optimize realistic 2010-2050 decarbonization roadmaps.
	\item Identify high impact technologies.
        \item Identify potential transition bottlenecks.
        \item Help Japan's policymakers create timelines for R\&D investment and infrastructure development.
        \item Quantify system sensitivity to technology readiness.
	\item Predict impediments to strategically optimal technology deployment.
	\end{itemize}
\end{description}

\end{alertblock}
%
%\begin{block}{Collaborators}
%
%I\textsuperscript{2}CNER collaborators will include primarily the members of 
%the Energy Analysis Division (EAD), including those members at other 
%institutes, universities or industries with connections to the EAD.   
%\end{block}

%----------------------------------------------------------------------------------------
%	METHODOLOGY
%----------------------------------------------------------------------------------------

\begin{block}{Introduction}
Previous work has compared the impact of innovative energy technologies in 
various world regions using \textbf{static} scenario analyses 
\cite{kikuchi_simulation-based_2017,pambudi_impact_2017}.  
We are simulating \textbf{dynamic} transition scenarios 
\cite{pfenninger_energy_2014} aimed at optimizing energy supply to minimize carbon 
emissions in Japan to 80\% of 1990 levels by 2050, and maintaining these levels from 2050-2100. \\

Our model includes conventional energy sources such as \textbf{fossil fuels, nuclear, wind, solar, hydropower, and geothermal}, and applicable I$^2$CNER technologies like \textbf{post-combustion absorption-type carbon capture and sequestration, hydrogen production (steam reforming and photocatalytic conversion)}. \hfill \break

\end{block}

%\vspace{77mm}

\begin{figure}[H] 
\centering
\includegraphics[scale=1.2]{conv_nuc_elc}
\caption{\textbf{Scenario 1} (conventional, with new nuclear) Electricity Generation.}
\label{s1e}
\end{figure}

\begin{figure}[H] 
\centering
\includegraphics[scale=1.2]{conv_nuc_co2}
\caption{\textbf{Scenario 1} (conventional, with new nuclear) CO$_2$ emissions.}
\label{s1c}
\end{figure}

\begin{figure}[H] 
\centering
\includegraphics[scale=1.2]{conv_nonuc_elc}
\caption{\textbf{Scenario 2} (conventional, no new nuclear) Electricity Generation.}
\label{s2e}
\end{figure}

\begin{figure}[H] 
\centering
\includegraphics[scale=1.2]{conv_nonuc_co2}
\caption{\textbf{Scenario 2} (conventional, no new nuclear) CO$_2$ emissions.}
\label{s2c}
\end{figure}

%----------------------------------------------------------------------------------------

\end{column} % End of the first column

%===========================================================================================
% SECOND COLUMN BEGINS
%=============================================================================================

%----------------------------------------------------------------------------------------

\begin{column}{\onecolwid} % The second column

%----------------------------------------------------------------------------------------
\begin{block}{Methodology}
        \textbf{The Integrated MARKAL-EFOM System (TIMES)} model generator 
        \cite{loulou_documentation_2005} 
        \cite{seebregts_energy/environmental_2002} optimizes energy systems 
        using linear and mixed-linear algorithms. A user-defined objective 
        function is solved within user defined constraints.\\
\textbf{Key assumptions:}

\begin{itemize}
        \item\textbf{Objective function:} System cost.

         \item\textbf{Constraints:} CO$_2$ emissions, demand (+1.7\% per year) \cite{noauthor_electricity_2017}.

         \item Hydropower and geothermal capacity held constant \cite{noauthor_energy_2018}.

         \item Variation in costs of wind, solar, and photocatalytic H$_2$ production is given in Fig. \ref{cost}.

         \item All new nuclear reactors are ABWRs \cite{rothwell_real_2006}.

         \item Growth rate and maximum annual increase in capacity of nuclear, wind and solar is based on data and trends seen in USA, China, and Japan. \cite{noauthor_electricity_2017,noauthor_energy_2018,eia_international_nodate,eia_monthly_2018,iea-pvps_snapshot_2018}

         \item \textbf{Carbon capture and sequestration and steam reforming:} Deployed in 2030, with its costs reducing by 2050 as per projected trends. \cite{kato_energy_2016}

         \item \textbf{Hydrogen production:} Steam reforming deployed in 2030. Photocatalytic hydrogen production deployed in 2050, assumed to be cost-competitive with other H$_2$ production technologies. \cite{kato_energy_2016} \cite{acar_comparative_2014} \\

 \end{itemize}

We have modeled \textbf{four scenarios} with different energy technologies enabled: \hfill \break

\begin{tabular}{c c c c c}

\textbf{Scenario} & \textbf{Figures}         & \textbf{Conventional}      & \textbf{I$^2$CNER}   & \textbf{New nuclear} \\
                  &                          & \textbf{technology}        &  \textbf{technology} & \textbf{reactors}    \\
1                 & Fig. \ref{s1e}\&\ref{s1c}&      \greencheck           &         \xmark       &      \greencheck     \\ 
2                 & Fig. \ref{s2e}\&\ref{s2c}&      \greencheck           &         \xmark       &         \xmark       \\ 
3                 & Fig. \ref{s3e}\&\ref{s3c}&      \greencheck           &      \greencheck     &      \greencheck     \\ 
4                 & Fig. \ref{s4e}\&\ref{s4c}&      \greencheck           &      \greencheck     &         \xmark       \\
\end{tabular}

\end{block}

%\vspace{65 mm}

\begin{figure}[H] 
\centering
\includegraphics[scale=1.2]{legend}
%\caption{Legend}
\label{legend}
\end{figure}

\begin{figure}[H] 
\centering
\includegraphics[scale=1.2]{i2cner_nuc_elc}
\caption{\textbf{Scenario 3} (with I$^2$CNER tech and new nuclear) Electricity Generation.}
\label{s3e}
\end{figure}

\begin{figure}[H] 
\centering
\includegraphics[scale=1.2]{i2cner_nuc_co2}
\caption{\textbf{Scenario 3} (with I$^2$CNER tech and new nuclear) CO$_2$ emissions.}
\label{s3c}
\end{figure}

\begin{figure}[H] 
\centering
\includegraphics[scale=1.2]{i2cner_nonuc_elc}
\caption{\textbf{Scenario 4} (with I$^2$CNER tech, no new nuclear) Electricity Generation.}
\label{s4e}
\end{figure}

\begin{figure}[H] 
\centering
\includegraphics[scale=1.2]{i2cner_nonuc_co2}
\caption{\textbf{Scenario 4} (with I$^2$CNER tech, no new nuclear) CO$_2$ emissions.}
\label{s4c}
\end{figure}

\end{column} % End of column 2, top half
\end{columns}
%=======================================================
% Double Wide Figure Column 
%=========================================================
%\input{model-fig}

\end{column} % End of wide column, total length

\begin{column}{\onecolwid} % The third column

%----------------------------------------------------------------------------------------
%	TAKE AWAY
%----------------------------------------------------------------------------------------
\begin{figure}[H] 
\centering
\includegraphics[scale=1.2]{cost}
\caption{Cost variation in solar, wind, and photocatalytic H$_2$ production.}
\label{cost}
\end{figure}
         \begin{alertblock}{Take Aways}
	\begin{itemize}
                \item Nuclear power has the lowest emissions intensity and it results in the most economical energy supply.

                \item Without nuclear and I$^2$CNER technology, solar and offshore wind must be deployed in a 1:2 ratio for grid resilience.

                \item Photocatalytic H$_2$ and offshore wind can meet I$^2$CNER targets in the absence of nuclear, but at significantly higher cost.

                \item Technologies with the highest impact so far: nuclear power, offshore wind, photocatalytic H$_2$ generation, solar power
	\end{itemize}
        \end{alertblock}

%--------------------------------------------------------
% PROJECT TIMELINE
%---------------------------------------------------------
\begin{block}{Limitations and Future work}
    \begin{itemize}

    \item Model focuses entirely on \textbf{electricity supply sector}.

    \item The cost of fossil fuels, geothermal, hydropower, steam reforming and nuclear is held constant.

    \item No limit on penetration of solar or wind except maximum potential.

    \item Annual timesteps and annually averaged capacity factors limit ability to capture seasonal and daily variation in wind and solar.

    \item Capture realistic, gradual transition behaviour and fluctuations in demand due to population decline in Japan.

    \item Incorporation of more I$^2$CNER technologies.

    \item Sensitivity analysis to maximize impact of I$^2$CNER technology.

    \item Detailed cost analysis

    \item Incorporation of natural gas and/or energy storage technologies to supplement renewable energy and I$^2$CNER technology.

    \end{itemize}
\end{block}

\begin{block}{References}
        {\footnotesize\bibliographystyle{ieeetr} 
        \bibliography{2019-chaube-i2cner-poster}}
\end{block}

%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\setbeamercolor{block alerted title}{fg=black,bg=norange} % Change the alert block title colors
\setbeamercolor{block alerted body}{fg=black,bg=white} % Change the alert block body colors

\begin{alertblock}{Contact Information}
\begin{itemize}
	\item Web: \href{arfc.github.io}{arfc.github.io}
	\item Email: \href{mailto:kdhuff@illinois.edu}{kdhuff@illinois.edu}
\end{itemize}

\end{alertblock}

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\setbeamercolor{block title}{fg=norange,bg=white} % Change the block title color

\begin{block}{Acknowledgements}

This research is being performed using funding received
from the International Institute for Carbon Neutral Energy Research (I$^2$CNER) 
Initiative on Challenges in Energy Assessment and Energy Transitions at  the  
University of Illinois under Director Petros Sofronis.

\vspace{10mm}
\begin{center}
\begin{tabular}{cccc}
\includegraphics[scale=0.5]{arfc_logo.png} & \includegraphics[scale=0.5]{wpi_logo.png} & \includegraphics[scale=0.1]{ku_logo.png} &\includegraphics[scale=0.7]{i2cner_logo.png}
%\includegraphics[width=\linewidth, height=0.1\textheight]{i2cner_logo.png}
\end{tabular}
\end{center}

\end{block}

\end{column} % End of the final column

\end{columns}
\end{frame} % End of the enclosing frame

\end{document}