### 2019-01-16 doc/2019-01-poster/presentation/cat_appearance.tex
added presentation template

\begin{frame}
  \frametitle{Columns}
  % a comment
        \begin{columns}
                \column[t]{5cm}
                Sometimes things need to be put side by side, in two nice 
                looking columns.

                Maybe one column involves a quotation.

                \begin{quote}
                        Explicit is better than implicit. -- The Zen of Python
                \end{quote}

                And, also, perhaps, a logo.
                \begin{center}
                        \includegraphics[height=0.2\textheight]{./images/arfc-logo}
                \end{center}
                \column[t]{5cm}
        \begin{figure}[htbp!]
        \begin{center}
      \includegraphics[height=4cm]{./images/kitten}
    \end{center}
          \caption{A caption describing the image. \cite{lastname_firstword_1900}.}
    \label{fig:kittenfigure}
  \end{figure}
        \end{columns}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Some Code}
        I have to use the fragile syntax for code slides.
\end{frame}