### 2019-01-17 doc/2019-01-poster/presentation/s4.tex

\begin{frame}
  \frametitle{Limitations}
  % a comment
        \begin{itemize}

         \item Focus entirely on \textbf{electricity supply sector}

         \item Cost of fossil fuels, geothermal, hydropower, steam reforming and nuclear is constant.

         \item Emissions tied to energy production, not capacity installation

         \item Annual time-step cannot capture seasonal/daily variation in wind, solar.

        \end{itemize}

\end{frame}