### 2019-01-17 doc/2019-01-poster/presentation/s4.tex
added presentation draft
Former-commit-id: 7b63d4afc4dc14f5c20f181f6611f6bb021750b2 [formerly 4e8b928e5f507f03cb62df0b8713b5f3a49f388b]
Former-commit-id: 2925c2287f07f3061567018c055b823d0fbff28f

\begin{frame}
  \frametitle{Scenario 4}
  % a comment
  \begin{figure}[htbp!]
    \begin{center}
      \includegraphics[scale=0.6]{./images/i2cner_nonuc_elc}
    \end{center}
          \caption{Scenario 4 Electricity Generation.}
    \label{s4e}
  \end{figure}
\end{frame}

\begin{frame}
  \begin{figure}[htbp!]
    \begin{center}
      \includegraphics[scale=0.6]{./images/i2cner_nonuc_co2}
    \end{center}
          \caption{Scenario 4 CO$_2$ emissions.}
    \label{s4c}
  \end{figure}

\end{frame}