### 2019-01-22 doc/2019-02-presentation/assums.tex
reorganized docs, corrected assumptions slide
Former-commit-id: f44c5c32bcf6c2f8dbb3642d31b2df8f8ad6ea89 [formerly 92ce9c2ab44f4d38b98b22cbd6208c607a6980d3]
Former-commit-id: edcc5a8790eae756807361a8b7d9692cf07d9259

\begin{frame}
  \frametitle{Assumptions}

  \begin{itemize}

  \item New Nuclear = ABWRs

  \item Nuclear, wind, solar growth rates based on trends in USA, China, Japan.

  \item CCS: Available for deployment in 2030, costs reduce by 2050.

  \item H$_2$: Steam reforming available in 2030, photocatalytic conversion in 2050.

  \item Hydropower, geothermal capacity held constant.

  \end{itemize}

\end{frame}