### 2019-01-23 doc/2019-02-presentation/basics.tex
minor punctuation corrections, added scenario frame title
Former-commit-id: 9e89b78dcdf4c533730cf1c055de182db6beb53b [formerly 20d17d0febcd06be3f506b362ab97d808910e4cf]
Former-commit-id: 02093b75aa75bd3201815b77b46f7e57b644889c

\begin{frame}
  \frametitle{Basics}
  % a comment
        \begin{itemize}

        \item Constrained linear/mixed-linear optimization problem.

        \item Objective function - System Cost.

        \item Constraints - CO$_2$ emissions, demand.

        \item Novel tech - H$_2$ (steam reforming, photocatalytic conversion), carbon capture and sequestration (absorption type).

        \end{itemize}
\end{frame}