### 2019-01-23 doc/2019-02-presentation/limits.tex

\begin{frame}
  \frametitle{Limitations}
  % a comment
        \begin{itemize}

         \item Focus entirely on \textbf{electricity supply sector}.

         \item Cost of fossil fuels, geothermal, hydropower, steam reforming and nuclear is constant.

         \item Emissions tied to energy production, not capacity installation.

         \item Annual time-step cannot capture seasonal/daily variation in wind, solar (however annually averaged availability factors are incorporated).

        \end{itemize}

\end{frame}