### 2019-01-24 doc/2019-02-presentation/basics.tex

\begin{frame}
  \frametitle{Basics}
  % a comment
        \begin{itemize}

        \item Constrained linear/mixed-linear optimization problem.

        \item Objective function - System Cost.

        \item Constraints - CO$_2$ emissions, demand.

        \item Novel tech - H$_2$ (steam reforming, photocatalytic conversion), absorption type carbon capture and sequestration (CCS).

        \end{itemize}
\end{frame}