### 2019-01-24 doc/2019-02-presentation/future.tex

\begin{frame}
  \frametitle{Future work}
   \begin{itemize}

   \item Fix premature retirement bug.

   \item Develop a scenario with more restricted nuclear (similar to METI projections).

   \item Transition from model creation to model refinement.

   \item Capture realistic, gradual transitions.

   \item Incorporate fluctuations in demand.

   \item Add energy storage to supplement renewables and H$_2$.

   \item Incorporate more I$^2$CNER technologies.

   \item Conduct sensitivity and cost analysis.

   \end{itemize}

\end{frame}