### 2019-01-25 doc/2019-02-presentation/basics.tex
added references, readjusted plot size to include caption, added c2peak factor table

\begin{frame}
  \frametitle{Basics}
  % a comment
        \begin{itemize}

        \item Constrained linear/mixed-linear optimization problem \cite{loulou_etsap-tiam:_2008}.

        \item Objective function - System Cost.

        \item Constraints - CO$_2$ emissions, demand.

        \item Novel tech - H$_2$ (steam reforming, photocatalytic conversion), absorption type carbon capture and sequestration (CCS)\cite{kato_energy_2016}.

        \end{itemize}
\end{frame}

%\begin{frame}
%\include{model-fig}
%\end{frame}