### 2019-01-26 doc/2019-02-presentation/conclusion.tex
fixes minor formatting issues, closes #73

\begin{frame}
  \frametitle{Conclusions}
        \begin{itemize}

        \item Highest impact (so far): nuclear, offshore wind, photocatalytic H$_2$, solar.

        \item Transition scenario cost (from \textbf{least} to \textbf{most} expensive):\\

        \begin{enumerate}

        \item Scenario 1 (conventional with nuclear) = Scenario 3 (I$^2$CNER with nuclear)

        \item Scenario 2 (conventional without nuclear)

        \item Scenario 4 (I$^2$CNER without nuclear)

        \end{enumerate}         
        \end{enumerate}        
%        Scenario 1(conventional with nuclear)=Scenario 3(I$^2$CNER with nuclear) $<$ \\
%        Scenario 2(conventional without nuclear) $<$ Scenario 4(I$^2$CNER without nuclear)

        \item H$_2$ and wind can meet I$^2$CNER goals at high cost without nuclear.

        \item Nuclear is the cleanest, cheapest source of energy.

        \item Without novel technology, solar and wind are deployed in a 1:2 ratio.

        \end{itemize}
\end{frame}