### 2019-05-08 doc/2019-05-report/acros.tex

%\newacronym{<++>}{<++>}{<++>}
%\newacronym{<++>}{<++>}{<++>}
\newacronym[longplural={metric tons of heavy metal}]{MTHM}{MTHM}{metric ton of heavy metal}
\newacronym{EDMC}{EDMC}{Energy and Data Modelling Center}
\newacronym{FEPC}{FEPC}{Federation of Electric Power Companies of Japan}
\newacronym{EIA}{EIA}{Energy Information Administration}
\newacronym{TIMES}{TIMES}{The Integrated Markal-EFOM System}
\newacronym{I$^2$CNER}{I$^2$CNER}{International Institute for Carbon-Neutral Energy Research}
\newacronym{CCS}{CCS}{Carbon Capture and Sequestration}
\newacronym{JWPA}{JWPA}{Japan Wind Power Association}
\newacronym{LCoE}{LCoE}{Levelized Cost of Electricity}
\newacronym{ABWR}{ABWR}{Advanced Boiling Water Reactors}
\newacronym{MoE}{MoE}{Ministry of Environment}
\newacronym{METI}{METI}{Ministry of Environment}{Ministry of Economy, Trade and Industry}
\newacronym{DSCINV}{DSCINV}{Discrete Investment}
\newacronym{C2Peak}{C2Peak}{Contribution to Peak}
\newacronym{MaxCAP}{MaxCAP}{Maximum Capacity}
%\acro{H2O}[$\mathrm{H_2O}$]{water}