### 2020-03-27 doc/2020-03-paper/Appendix.tex
cleanup
Former-commit-id: 58322421ef92fff10ab24b0e4c52512d81771895
Former-commit-id: e387f21dccdfc3dd76f67082f9b0cf791fe07277

\section{Feed rates for all cases} \label{Appendix}
Appendix