### 2020-03-27 doc/2020-03-paper/acks.tex

\section{Acknowledgments}

Osama Ashraf would like to thank the Egyptian Ministry of Higher Education (MoHE), as well as MEPhI's Competitiveness Program for providing financial support for this research. The facility and tools needed to conduct this work were supported by MEPhI.

The authors contributed to this work as described below.

Osama Ashraf conceived and designed the simulations, wrote the paper, prepared figures 
and/or tables, performed the computation work, and reviewed drafts of the paper.

Andrei Rykhlevskii conceived and designed the simulations, wrote the paper, prepared figures 
and/or tables, performed the computation work, and reviewed drafts of the paper. Andrei Rykhlevskii 
is supported by  DOE ARPA-E MEITNER program award DE-AR0000983.

G. V. Tikhomirov directed and supervised the work, conceived and designed the simulations and reviewed drafts of the paper. Prof. Tikhomirov is Deputy Director of the Institute of Nuclear Physics and Engineering MEPhI. Board member of Nuclear society of Russia.

Kathryn D. Huff supervised the work, conceived and contributed to conception of the simulations, and reviewed drafts of the paper.  Prof. Huff is supported by the Nuclear Regulatory Commission Faculty Development Program, the National Center for Supercomputing Applications, the International Institute for Carbon Neutral Energy Research (WPI-I2CNER), 
sponsored by the Japanese Ministry of Education, Culture, Sports, Science and Technology, and  DOE ARPA-E MEITNER program award DE-AR0000983.

This research is part of the Blue Waters sustained-petascale computing 
project, which is supported by the National Science Foundation (awards 
OCI-0725070 and ACI-1238993) and the state of Illinois. Blue Waters is a joint 
effort of the University of Illinois at Urbana-Champaign and its National 
Center for Supercomputing Applications.

The authors thank members of the Advanced Reactors and Fuel Cycles (ARFC) 
group at the University of Illinois at Urbana-Champaign for helpful 
discussions relating to this paper. Finally, the authors would like to thank 
Matthew Kozak for constructive criticism of the manuscript.