### 2020-03-27 doc/2020-03-paper/conclusion.tex

\section{Conclusion} \label{Conclusion}
Conclusion
\section{Future work}
FUTURE WORK

\section{Declaration of Competing Interest}

NECESARY?