### 2020-04-06 doc/2020-03-paper/acros.tex
cleanup, added section 3 text
Former-commit-id: fcbb2d2478504439efa3afcf874bcf8fae6bd2f3
Former-commit-id: 7f203aacef92cb574a86b988059c9dad3d6e97ef

\newacronym{<++>}{<++>}{<++>}
\newacronym{I$^2$CNER}{I$^2$CNER}{International Institute for Carbon Neutral Energy Research}
\newacronym{MARKAL}{MARKAL}{MARKet ALlocation}
\newacronym{TIMES}{TIMES}{The Integrated MARKAL-EFOM System}
\newacronym{EDMC}{EDMC}{The Energy Data and Modelling Centre}
\newacronym{IEEJ}{IEEJ}{The Institute of Energy Economics, Japan}
\newacronym{CCS}{CCS}{Carbon Capture and Sequestration}
\newacronym{USC}{USC}{Ultra Supercritical Coal}
\newacronym{OM}{O\&M}{Operation and Maintenance Cost}
%\newacronym[longplural={metric tons of heavy metal}]{MTHM}{MTHM}{metric ton of heavy metal}