### 2020-04-07 doc/2020-03-paper/acks.tex
updates to paper
Former-commit-id: 1fff218c908a2ec1610b116ab213625083a8c178
Former-commit-id: 649db3cb4c71aa18e09ad8a46da4b1e629c4efa6

\section{Acknowledgments}

The authors contributed to this work as described below.

Osama Ashraf conceived and designed the simulations, wrote the paper, prepared figures 
and/or tables, performed the computation work, and reviewed drafts of the paper.

Andrei Rykhlevskii conceived and designed the simulations, wrote the paper, prepared figures 
and/or tables, performed the computation work, and reviewed drafts of the paper. Andrei Rykhlevskii 
is supported by  DOE ARPA-E MEITNER program award DE-AR0000983.

Kathryn D. Huff supervised the work, conceived and contributed to conception of the simulations, and reviewed drafts of the paper.  Prof. Huff is supported by the Nuclear Regulatory Commission Faculty Development Program, the National Center for Supercomputing Applications, the International Institute for Carbon Neutral Energy Research (WPI-I2CNER), 
sponsored by the Japanese Ministry of Education, Culture, Sports, Science and Technology, and  DOE ARPA-E MEITNER program award DE-AR0000983.

This research is part of the Blue Waters sustained-petascale computing 
project, which is supported by the National Science Foundation (awards 
OCI-0725070 and ACI-1238993) and the state of Illinois. Blue Waters is a joint 
effort of the University of Illinois at Urbana-Champaign and its National 
Center for Supercomputing Applications.

The authors thank members of the Advanced Reactors and Fuel Cycles (ARFC) 
group at the University of Illinois at Urbana-Champaign for helpful 
discussions relating to this paper. Finally, the authors would like to thank 
Matthew Kozak for constructive criticism of the manuscript.