### 2020-06-07 doc/2020-03-paper/acros.tex
added data + references
Former-commit-id: d21d3fa43c141a146410ddc02c412e9aeb411b66
Former-commit-id: 5dd078789f042d8cb1db3ad9cced13317e75feb5

\newacronym{<++>}{<++>}{<++>}
\newacronym{I$^2$CNER}{I$^2$CNER}{International Institute for Carbon Neutral Energy Research}
\newacronym{MARKAL}{MARKAL}{MARKet ALlocation}
\newacronym{TIMES}{TIMES}{The Integrated MARKAL-EFOM System}
\newacronym{EDMC}{EDMC}{The Energy Data and Modelling Centre}
\newacronym{IEEJ}{IEEJ}{The Institute of Energy Economics, Japan}
\newacronym{CCS}{CCS}{Carbon Capture and Sequestration}
\newacronym{USC}{USC}{Ultra Supercritical Coal}
\newacronym{OM}{O\&M}{Operation and Maintenance Cost}
\newacronym{PEMFC}{PEMFC}{Polymer Electrolyte Membrane Fuel Cells}
\newacronym{lcoe}{LCOE}{Levelized Cost of Electricity}
\newacronym{lng}{LNG}{Liquified Natural Gas}
\newacronym{AEC}{AEC}{Alkaline Electrolyzer Cell}
\newacronym{PEMEC}{PEMEC}{Polymer Electrolyte Membrane Electrolyzer Cell}
\newacronym{SOEC}{SOEC}{Solid Oxide Electrolyzer Cell}
\newacronym{SOFC}{SOFC}{Solid Oxide Fuel Cell}
\newacronym{PWS}{PWS}{Photochemical Water Splitting}
%\newacronym[longplural={metric tons of heavy metal}]{MTHM}{MTHM}{metric ton of heavy metal}