### 2020-06-22 doc/2020-03-paper/acks.tex
added plots for base scen 1-4
Former-commit-id: 3d0d846c0d9516d8f3e77a5968cef48b21b24b62
Former-commit-id: b13122957eda23ef887beeafd327aaf3b6cddf4d

\section{Acknowledgments}

The authors contributed to this work as described below.

Kathryn D. Huff supervised the work, conceived and contributed to conception of the simulations, and reviewed drafts of the paper.  Prof. Huff is supported by the Nuclear Regulatory Commission Faculty Development Program, the National Center for Supercomputing Applications, the International Institute for Carbon Neutral Energy Research (WPI-I2CNER), 
sponsored by the Japanese Ministry of Education, Culture, Sports, Science and Technology, and  DOE ARPA-E MEITNER program award DE-AR0000983.