### 2020-06-22 doc/2020-03-paper/discussion.tex
added results
Former-commit-id: 4c3b33b13ba2639f7649dce115d0832e5f3badac
Former-commit-id: 5c0328da46f31d8f48794b2dba882d21e6c30d7b

\section{Discussion}
% This should explore the significance of the results of the work, not repeat
% them. A combined Results and Discussion section is often appropriate. Avoid
% extensive citations and discussion of published literature.