### 2020-06-23 doc/2020-03-paper/acks.tex
section 3 added parameter details, added section 5 notes
Former-commit-id: 3c3e5fdaddb8f4d18f3fe5b84bec77119d161dcf
Former-commit-id: fd5392759009fc1b04c95f5123799685cb74ee9b

\section{Acknowledgments}

%The authors contributed to this work as described below.
%
%
%Kathryn D. Huff supervised the work, conceived and contributed to conception of the simulations, and reviewed drafts of the paper.  Prof. Huff is supported by the Nuclear Regulatory Commission Faculty Development Program, the National Center for Supercomputing Applications, the International Institute for Carbon Neutral Energy Research (WPI-I2CNER), 
%sponsored by the Japanese Ministry of Education, Culture, Sports, Science and Technology, and  DOE ARPA-E MEITNER program award DE-AR0000983.