### 2020-06-23 doc/2020-03-paper/conclusion.tex
added results details
Former-commit-id: ed19f4a90ca89ba623e169f45f26da001cb413db
Former-commit-id: d18cad00c2db357dbb0f67988ccf31fec6140bba

\section{Conclusion} \label{Conclusion}

\section{Future work}

\section{Declaration of Competing Interest}