### 2020-09-16 doc/2020-03-paper/discussion.tex

\section{Discussion}
% This should explore the significance of the results of the work, not repeat
% them. A combined Results and Discussion section is often appropriate. Avoid
% extensive citations and discussion of published literature.
%Notes for Ansh:
%Importance of nuclear : cost of various transitions, flexibility by itself and when coupled with storage (Li-ion, H2). How it precludes the need for CCS.
%Urgency of transition : scenarios + sensitivity analysis.
%Li-ion storage: land use requirements for each case, qualitatively discuss Co, Ni requirements.
%AECs,PEMECs,SOECs need to be highly flexible with high availability factors.
%SOFCs: need for extremely responsive SOFCs with high availability factor, flexible operation (cite challenges from NREL).