### 2020-09-18 doc/2020-03-paper/conclusion.tex
adds clean results for syscost, net emissions
Former-commit-id: 5a0d7dd093d146c65fbf51a7f4887799ed403066
Former-commit-id: 8d484390baa9c26c0788b9cfa933ef94f0c1e0fc

\section{Conclusion} \label{Conclusion}
%what we did. sum up key technologies, what needs to happen for technologies that are not "key", importance of rapid change/suggested timeline.

\section{Future work}
%grid resilience - analysis of grid stability of similar mixes at the resolution of minutes. Maybe in TIMES (ugh).
\section{Declaration of Competing Interest}