### 2020-12-09 doc/2020-03-paper/introduction.tex

\section{Introduction} \label{Introduction}
In order to mitigate climate change, and to improve environmental outcomes, many nations are actively seeking to reduce carbon emissions, and have formalized this goal through the Paris Agreements \cite{united_nations_framework_convention_on_climate_change_unfccc_submission_2015}. The largest contribution to global \gls{GHG} emissions, some 73\%, comes from largely from energy consumption of the transportation, electricity and heat, buildings, manufacturing, and construction sectors \cite{ge_4_2020}. For developed economy nations rich in natural resources, switching to natural gas or implementing \gls{CCS} on fossil fuel plants is one option to reduce GHG emissions. However these options are not always economically feasible for developing nations, which will lead to increased emissions through greater coal use \cite{international_energy_agency_latest_2019}. For Japan, a developed nation without fossil fuel resources, the challenge to reduce GHG emissions is likely to follow a different path, already evidenced by the restart of nuclear reactors and a shift toward large-scale renewable energy deployment \cite{international_energy_agency_latest_2019}.

Although influenced by the Paris Agreements, Japanese energy policy is governed by the Basic Energy Plan, recently updated to the 5th edition and approved by the Cabinet in July of 2018 \cite{noauthor_japans_2018}. The Basic Energy Plan outlines national policy toward a new energy system for the years 2030 and 2050, cognizant of limited indigenous resources, the impact of the Fukushima incident, and external pressures on energy supplies \cite{meti_annual_2018}. The plan reaffirms the Japanese benchmarks for evaluating the energy system, first and foremost, within the context of energy security, followed by economic efficiency, safety, and a consideration of the environment (summarized as '3E+S'; ibid). Although there are some parallels between the 3E+S goals and the Paris Agreement targets agreed to by Japan, the plan does not detail how the 2050 emission reduction target of 80\% is to be met. This vagueness may be due to uncertainty regarding the economical feasibility of GHG emissions reduction in each sector. It has been suggested that electrification of a number of sectors will be required to achieve the ambitious 2050 target, underpinned by low-carbon technologies \cite{matsuo_quantitative_2018}. For the power sector to achieve such a target, near-zero emissions are required, and early action utilizing existing technologies is preferable to delayed action in preference for future technologies \cite{ashina_roadmap_2012}. It is likely that a mixture of current and emerging technologies will be employed to achieve carbon reduction targets in Japan. The strategies currently under consideration include reinvigorating nuclear power, deploying \gls{CCS} to fossil fuel power plants, and ushering in the hydrogen economy based on renewable energy-based electrolysis as well as hydrogen imports from abroad \cite{ashina_roadmap_2012, matsuo_quantitative_2018, noauthor_basic_2017}.

The aim of this research is to investigate the likely suite of electricity generation and storage technologies, as well as their feasibility in meeting Japan's carbon reduction goals, while being cognizant of energy policy, resource limitations, demand growth, emerging technologies, and economic constraints using \gls{TIMES} framework. Our dynamic simulations of transition scenarios, which focus on minimizing the cost of the transition while satisfying CO$_2$ emission constraints, suggest potential economically feasible decarbonization pathways that meet the increasing near-term electricity demand. Additionally, the significance of key economic parameters of emerging technologies is assessed through sensitivity analysis, in order to highlight the most impactful parameters and hence guide research and development efforts focused on these technologies.