### 2020-12-10 doc/2020-03-paper/acros.tex

\newacronym{<++>}{<++>}{<++>}
\newacronym{I$^2$CNER}{I$^2$CNER}{International Institute for Carbon Neutral Energy Research}
\newacronym{MARKAL}{MARKAL}{MARKet ALlocation}
\newacronym{TIMES}{TIMES}{The Integrated MARKAL-EFOM System}
%\newacronym{EDMC}{EDMC}{The Energy Data and Modelling Centre}
\newacronym{IEEJ}{IEEJ}{The Institute of Energy Economics, Japan}
\newacronym{CCS}{CCS}{carbon capture and sequestration}
\newacronym{USC}{USC}{ultra-supercritical coal}
\newacronym{OM}{O\&M}{operation and maintenance}
\newacronym{PEMFC}{PEMFC}{polymer electrolyte membrane fuel cell}
\newacronym{lcoe}{LCOE}{levelized cost of electricity}
\newacronym{lng}{LNG}{liquified natural gas}
\newacronym{AEC}{AEC}{alkaline electrolyser cell}
\newacronym{PEMEC}{PEMEC}{polymer electrolyte membrane electrolyser cell}
\newacronym{SOEC}{SOEC}{solid oxide electrolyser cell}
\newacronym{SOFC}{SOFC}{solid oxide fuel cell}
\newacronym{PWS}{PWS}{photochemical water splitting}
\newacronym{INDC}{INDC}{Intended Nationally Determined Contribution}
\newacronym{GHG}{GHG}{greenhouse gas}
\newacronym{BOP}{BOP}{balance of plant}
\newacronym{P2G}{P2G}{power-to-gas}
\newacronym{FCV}{FCV}{fuel cell vehicle}
%\newacronym[longplural={metric tons of heavy metal}]{MTHM}{MTHM}{metric ton of heavy metal}