### 2020-12-14 publications/2020-03-paper/acks.tex

\section{Acknowledgments}

The author(s) gratefully acknowledge the support of the International Institute for Carbon Neutral Energy Research (WPI-I2CNER), sponsored by the Japanese Ministry of Education, Culture, Sports, Science and Technology. We are also grateful to Kenshi Itaoka and Hadi Farabi-Asl for providing valuable discussions related to this project, and to Gavin Davis and Nataly Panczyk for editing drafts.

The authors contributed to this work as described below.

Anshuman Chaube designed simulation models, conducted sensitivity analysis, and wrote the original draft of the paper.

Andrew Chapman conceived and contributed to conception of the simulations, designed simulation models, and wrote the original draft of the paper.

Akari Minami translated data from Japanese to English, and assisted with testing and performing simulations.

James Stubbins supervised the work, conceived and contributed to conception of the simulations, and reviewed drafts of the paper.

Kathryn D. Huff supervised the work, conceived and contributed to conception of the simulations, and reviewed drafts of the paper.  Prof. Huff is supported by the Nuclear Regulatory Commission Faculty Development Program, the National Center for Supercomputing Applications, the International Institute for Carbon Neutral Energy Research (WPI-I$^2$CNER), sponsored by the Japanese Ministry of Education, Culture, Sports, Science and Technology, and  DOE ARPA-E MEITNER program award DE-AR0000983.