### 2021-01-13 publications/2020-03-paper/cover-letter/cover_letter.tex
adds cover letter draft
Former-commit-id: 51cf05d2f0977e5649586ad8d0ae4866b0007132
Former-commit-id: 8ed31854fbef04f7ca7e0967c806c5db6d33e6ae

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%       PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt]{letter} % Default font size of the document, change to 10pt to fit more text
%\usepackage{graphicx}
%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\usepackage[left=1.25in,right=1.25in,top=1in,bottom=0.5in]{geometry}
%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right
%\usepackage{everypage}
%\usepackage{blindtext}
%\usepackage{eso-pic,graphicx}
 \begin{document}
%\AddToShipoutPictureBG*{\includegraphics[width=\paperwidth,height=\paperheight]{background.pdf}}

%----------------------------------------------------------------------------------------
%       ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professor Jinyue Yan\\
Editor-in-Chief, Applied Energy.}

%----------------------------------------------------------------------------------------
%       YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

\address{Anshuman Chaube\\
achaube2@illinois.edu\\
226 Talbot Laboratory\\
MC-234\\
104 S. Wright Street\\
Urbana, IL 61801}

%----------------------------------------------------------------------------------------
%       LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{Dear Professor Yan,}

Please find enclosed our manuscript entitled: ``The role of current and emerging technologies in meeting Japan's mid- to long-term carbon reduction goals for Japan'' which I and my co-authors, Chapman, Minami, Stubbins, and Huff, 
are submitting for exclusive consideration for publication as a research 
article in Progress in Nuclear Energy.

Our work uses TIMES framework to explore decarbonisation pathways for the Japanese electricity supply sector using conventional and emerging (hydrogen, carbon capture and sequestration, and emerging photovoltaic cells) technologies. This work is novel in its inclusion of nascent technologies like solid oxide electrolysers and photochemical water splitting reactors, long-term impact evaluation up to the year 2100, life-cycle emissions for all technologies, and projections and learning curves for parameters like capital costs. We assess novel technologies’ impact using a rigorous sensitivity analysis.

Our results approach true carbon-neutrality with the incorporation of lifecycle emissions. We also find that Japan can achieve energy security while mitigating emissions using renewable-based hydrogen. Further, incorporation of nuclear along with hydrogen emerges as an economically viable strategy with continued emission reduction until 2100. Our sensitivity analysis identifies key emerging technologies, and their most impactful parameters.

Our work will offer a detailed comparative assessment of transition strategies to global energy analysts and Japanese policymakers. Our sensitivity analysis will also inform near-term deployment goals of mature technologies and long-term research objectives for lab-scale technologies in Japan.

We confirm that we have no conflicts of interest to disclose, and that this work has been reviewed by native speakers of English (Chapman, Stubbins, and Huff). I acknowledge my availability to review 3 or more articles for Applied Energy. Please direct your correspondence apropos this manuscript to me at achaube2@illinois.edu.

\closing{Sincere regards,
%\includegraphics[height=2cm]{signature}\\
\fromsig{Anshuman Chaube\\PhD Candidate\\Nuclear, Plasma, \& Radiological 
Eng.\\University of Illinois\\achaube2@illinois.edu}
}

%----------------------------------------------------------------------------------------

\end{letter}

\end{document}