### 2021-01-15 publications/2020-03-paper/abstract.tex
adds kdh corretions, glossaries, CRediT statement
Former-commit-id: 63eaba071d907d33887c9080bc221f79aae32636
Former-commit-id: 1b7ddd720025f2dbe7f1ec4b96fdcca1d1df9dd9

\begin{abstract}

We simulated possible pathways to meeting 2030 and 2050 emission targets within the Japanese electricity supply sector using a single-region \gls{TIMES} model. Critically, our simulations incorporate novel technologies like hydrogen electrolysers, carbon capture, photochemical water splitting, and emerging photovoltaic cells, assess long-term impact up to the year 2100, and include life-cycle emissions and learning curves for parameters such as investment cost, efficiency, and emission coefficients. Results indicate that a hybrid approach, using nuclear power and hydrogen from renewable energy-based electrolysis, is cost-effective and provides long-term emission reduction along with energy security. Nuclear, wind, solar, and hydrogen from renewables emerge as key emission reduction technologies, while natural gas with carbon capture plays a minor role.

\end{abstract}