### 2021-01-15 publications/2020-03-paper/acks.tex

\section{CRediT Statement}

The authors contributed to this work as described below.

\textbf{Anshuman Chaube}: Methodology, validation, writing - original draft, visualisation.

\textbf{Andrew Chapman}: Methodology, validation,  writing - original draft, writing - review and editing.

\textbf{Akari Minami}: data curation, validation.

\textbf{James Stubbins}: writing - review and editing, supervision.

\textbf{Kathryn D. Huff}: Conceptualisation, resources, writing - review and editing, supervision.

\section{Acknowledgments}

The author(s) gratefully acknowledge the support of the International Institute for Carbon Neutral Energy Research (WPI-I2CNER), sponsored by the Japanese Ministry of Education, Culture, Sports, Science and Technology.

Prof. Huff is supported by the Nuclear Regulatory Commission Faculty Development Program, the National Center for Supercomputing Applications, the International Institute for Carbon Neutral Energy Research (WPI-I$^2$CNER), sponsored by the Japanese Ministry of Education, Culture, Sports, Science and Technology, and  DOE ARPA-E MEITNER program award DE-AR0000983.

We are also grateful to Kenshi Itaoka and Hadi Farabi-Asl for providing valuable discussions related to this project, and to Gavin Davis and Nataly Panczyk for editing drafts.