### 2021-01-15 publications/2020-03-paper/highlights/highlights.tex
adds highlights
Former-commit-id: d81cc9b7b4e78b95d3602c15c5d80f5297fae011
Former-commit-id: d4cf835837583cac16171310086eda753fa38615

\documentclass[review]{elsarticle}

\usepackage{lineno}
\usepackage{xspace}
\modulolinenumbers[5]

\journal{Applied Energy}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

%%%% packages and definitions (optional)
\usepackage{placeins}
\usepackage{booktabs} % nice rules (thick lines) for tables
\usepackage{microtype} % improves typography for PDF
\usepackage{hhline}
\usepackage{amsmath}

%\usepackage[demo]{graphicx}
%\usepackage{caption}
%\usepackage{subcaption}

\usepackage{booktabs}
\usepackage{threeparttable, tablefootnote}

\usepackage{tabularx}
\newcolumntype{b}{>{\hsize=1.0\hsize}X}
\newcolumntype{s}{>{\hsize=.5\hsize}X}
\newcolumntype{m}{>{\hsize=.75\hsize}X}
\newcolumntype{x}{>{\hsize=.25\hsize}X}

\graphicspath{ {figures/} }

% tikz %
\usepackage{tikz}
\usetikzlibrary{positioning, arrows, decorations, shapes}

\usetikzlibrary{shapes.geometric,arrows}
\tikzstyle{process} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30]
\tikzstyle{object} = [ellipse, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=green!30]
\tikzstyle{arrow} = [thick,->,>=stealth]

% hyperref %
\usepackage[hidelinks]{hyperref}

\begin{document}
\begin{frontmatter}
\title{The role of current and emerging technologies in meeting Japan's mid- to long-term carbon reduction goals}
\end{frontmatter}
\section*{Highlights}
\begin{itemize}
	\item Transition pathways using nuclear, renewables, carbon capture, and hydrogen.
	\item Includes long-term assessment up to 2100 
	\item Dynamic simulations use learning curves for cost and emissions..
	\item Sensitivity analysis identifies key parameters and thresholds.
\end{itemize}
\end{document}