### 2021-01-19 publications/2020-03-paper/abstract.tex
incorporates the last of kdh comments, closes #98 #99
Former-commit-id: a009126bcf29f293c8ef7aeda35d1402a6cfb4c2
Former-commit-id: bf3d6a2e2d208737784bc145af251c1a2db4774b

\begin{abstract}

We simulated possible pathways to meeting 2030 and 2050 emission targets within the Japanese electricity supply sector using a single-region model in \gls{TIMES}. Critically, our simulations incorporate novel technologies like hydrogen electrolysers, carbon capture, photochemical water splitting, and emerging photovoltaic cells, assess long-term impact up to the year 2100, and include life-cycle emissions and learning curves for parameters such as investment cost, efficiency, and emission coefficients. Results indicate that a hybrid approach, using nuclear power and hydrogen from renewable energy-based electrolysis, is cost-effective and provides long-term emission reduction along with energy security. Nuclear, wind, solar, and hydrogen from renewables emerge as key emission reduction technologies, while natural gas with carbon capture plays a minor role.

\end{abstract}