### 2021-02-02 publications/2020-03-paper/credit-statements/credit.tex
separates credit statement, declaration of interest based on applied energy guidelines
Former-commit-id: 15bdec437b737268ef40dcbb97979ba5d996b60a
Former-commit-id: c1132ae75c7fe5215b0927d3a86a69bc7c6786ca

\documentclass[review]{elsarticle}

\usepackage{lineno}
\usepackage{xspace}
\modulolinenumbers[5]

\journal{Applied Energy}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

%%%% packages and definitions (optional)
\usepackage{placeins}
\usepackage{booktabs} % nice rules (thick lines) for tables
\usepackage{microtype} % improves typography for PDF
\usepackage{hhline}
\usepackage{amsmath}

%\usepackage[demo]{graphicx}
%\usepackage{caption}
%\usepackage{subcaption}

\usepackage{booktabs}
\usepackage{threeparttable, tablefootnote}

\usepackage{tabularx}
\newcolumntype{b}{>{\hsize=1.0\hsize}X}
\newcolumntype{s}{>{\hsize=.5\hsize}X}
\newcolumntype{m}{>{\hsize=.75\hsize}X}
\newcolumntype{x}{>{\hsize=.25\hsize}X}

\graphicspath{ {figures/} }

% tikz %
\usepackage{tikz}
\usetikzlibrary{positioning, arrows, decorations, shapes}

\usetikzlibrary{shapes.geometric,arrows}
\tikzstyle{process} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30]
\tikzstyle{object} = [ellipse, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=green!30]
\tikzstyle{arrow} = [thick,->,>=stealth]

% hyperref %
\usepackage[hidelinks]{hyperref}

\begin{document}
\begin{frontmatter}
\title{The role of current and emerging technologies in meeting Japan's mid- to long-term carbon reduction goals}
\end{frontmatter}
\section*{CRediT Statement}

The authors contributed to this work as described below.

\textbf{Anshuman Chaube}: Methodology, validation, writing - original draft, visualisation.

\textbf{Andrew Chapman}: Methodology, validation,  writing - original draft, writing - review and editing.

\textbf{Akari Minami}: data curation, validation.

\textbf{James Stubbins}: writing - review and editing, supervision.

\textbf{Kathryn D. Huff}: Conceptualisation, resources, writing - review and editing, supervision.
\end{document}