### 2021-06-21 publications/2020-03-paper/revisions/cover-letter/cover_letter.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%       PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt]{letter} % Default font size of the document, change to 10pt to fit more text
%\usepackage{graphicx}
%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\usepackage[left=1.25in,right=1.25in,top=1in,bottom=0.5in]{geometry}
%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right
%\usepackage{everypage}
%\usepackage{blindtext}
\usepackage{eso-pic}
\usepackage{graphicx}

 \begin{document}
%\ClearShipoutPicture
\AddToShipoutPictureBG*{\includegraphics[width=\paperwidth,height=\paperheight]{background.pdf}}

%----------------------------------------------------------------------------------------
%       ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professor Jinyue Yan\\
Editor-in-Chief, Applied Energy.}

%----------------------------------------------------------------------------------------
%       YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

\address{Anshuman Chaube\\
achaube2@illinois.edu\\
226 Talbot Laboratory\\
MC-234\\
104 S. Wright Street\\
Urbana, IL 61801}

%----------------------------------------------------------------------------------------
%       LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{Dear Professor Yan,}

Please find enclosed a revised version of our manuscript entitled: ``The role of current and emerging technologies in meeting Japan's mid- to long-term carbon reduction goals'' which I and my co-authors, Chapman, Minami, Stubbins, and Huff 
are submitting for exclusive consideration for publication as a research article in Applied Energy.

Thank you for giving us the opportunity to revise and resubmit our manuscript. The comments provided by the reviewers were considered seriously and their feedback has improved the readability and academic rigour of our work.

We look forward to working with you towards moving towards publication.

We confirm that we have no conflicts of interest to disclose, and that this work has been reviewed by native speakers of English (Chapman, Stubbins, and Huff). I acknowledge my availability to review 3 or more articles for Applied Energy. Please direct your correspondence apropos this manuscript to me at achaube2@illinois.edu.

\closing{Sincere regards,\\
%\includegraphics[height=2cm]{signature}\\
\fromsig{Anshuman Chaube\\PhD Candidate\\Nuclear, Plasma, \& Radiological 
Eng.\\University of Illinois at Urbana-Champaign.\\achaube2@illinois.edu}
}

%----------------------------------------------------------------------------------------

\end{letter}

\end{document}